### 2020-06-11 ExperimentalData/DIIID/169466/Results/Report/Report.tex
Improved neutral model

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{graphicx}
\allowdisplaybreaks

\title{\bf DIII-D Discharge 169466}
\date{}
\author{}

\begin{document}
\maketitle

\section{Llama Data}
Figure~\ref{fig1} shows the following fit to the Llama data:
\begin{equation}
n_n(R) = \frac{n_n}{1+(R-R_{\rm sep})^{\,2}/L^{\,2}},
\end{equation}
where $n_n = 1.4\times 10^{17}\,{\rm m}^{-3}$, $L= 0.015\,{\rm m}$, $R_{\rm sep}= 1.92\,{\rm m}$.

\begin{figure}[h]
\centering
\includegraphics[height=4in]{neutral.eps}
\caption{Fit to Llama data. Black: Data points. Red: Fit.}\label{fig1}
\end{figure}

\section{Natural Frequencies}
\begin{figure}
\centering
\includegraphics[height=4in]{3200.eps}
\caption{Natural frequencies at $t=3200$ ms. Red: Linear natural frequency. Blue: Nonlinear natural frequency. Black: ${\bf E}\times {\bf B}$ frequency.}\label{fig2}
\end{figure}

\begin{figure}
\centering
\includegraphics[height=4in]{3600.eps}
\caption{Natural frequencies at $t=3600$ ms. Red: Linear natural frequency. Blue: Nonlinear natural frequency. Black: ${\bf E}\times {\bf B}$ frequency.}\label{fig3}
\end{figure}

\end{document}