### 2020-11-05 Papers/Paper2/Submit/Reply.tex

\documentclass{article}[12pt]
\usepackage{fullpage}

\begin{document}
\begin{center}
{\em Further Modeling $q_{95}$ Windows for the Suppression of Edge Localized Modes by Resonant Magnetic}\\[0.5ex]
{\em Perturbations in the DIII-D tokamak}\\[1ex]
by Richard Fitzpatrick\\[1ex]
{\bf Reply to Editor's Comments}
\end{center}

I have modified the paper to remove material that appeared in earlier papers. I have also tried to make it more
clear why it is somewhat surprising that the relaxation of the no-slip constraint makes so little difference to
the calculation.
\end{document}